---
title: "p8105_hw3_xw2598"
author: Xinyao Wu
date: 2018-10-05
output: github_document
---

##problem1

```{r}
#build a dataset
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
library(tidyverse)
library(ggridges)
data(brfss_smart2010)
brfss = brfss_smart2010 %>% 
#data cleaning
#rename variables
  janitor::clean_names() %>% 
#focus on the “Overall Health” topic
  filter(topic == "Overall Health") %>% 
#include only responses from “Excellent” to “Poor”  (1)???
  select(year,locationabbr,locationdesc,response,data_value) %>% 
#organize responses as a factor
  mutate(
    response = factor(response)
  )
head(brfss, 10)
```


```{r}
#In 2002, which states were observed at 7 locations
loc_brfss = brfss %>% 
  group_by(year,locationabbr) %>% 
  summarise(
    count = n_distinct(locationdesc)
    )
  filter(loc_brfss,count == 7,year == 2002)
```
Comments: In 2002,CT,FL,NC were observed at 7 locations. I choose distinct specific locations which were observed in each state as the dataset, because from the original data each specific location were repetitive and the repetity doesn`t make sence in this question.

```{r}
#spaghetti plot
ggplot(loc_brfss, aes(x = year, y = count, fill = locationabbr,color = locationabbr))+
  geom_line() +
  labs(
    title = "Observations number in each state from 2002 to 2010 plot",
    x = "year",
    y = "Obervation numer"
  )+
  scale_color_hue(name = " state", h = c(0,360))
#comments


```


@Using this dataset, do or answer the following (commenting on the results of each):
@Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

 
